---
title: "GAPMINDER | Happiness over the years"
format:
  dashboard:
    embed-resources: true
theme: lux
---
  
```{r}
# Load packages 
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, gapminder,
               bslib, bsicons, shiny, 
               rnaturalearth, plotly, 
               countrycode, htmltools, 
               reactable, here
)

```

```{r}
pop <- read_csv(here("data/pop.csv"))

pop_iso <- pop %>% 
  pivot_longer(-country, names_to = "year", values_to = "pop") %>% 
  mutate(year = as.integer(year)) %>% 
  mutate(pop = case_when(
    str_detect(pop, 'k$') ~ as.numeric(str_remove(pop, "k")) * 1000,
    str_detect(pop, 'M$') ~ as.numeric(str_remove(pop, "M")) * 1000000,
    str_detect(pop, 'B$') ~ as.numeric(str_remove(pop, "M")) * 1000000000,
    TRUE ~ as.numeric(pop))) %>% 
  mutate(country_iso = countrycode(country, "country.name", "iso3c"))

happy_data <- read_csv(here("data/hapiscore_whr.csv")) %>% 
  pivot_longer(`2005`:`2022`, names_to = "year", values_to = "happiness") %>% 
  mutate(year = as.integer(year)) %>% 
  mutate(country_iso = countrycode(country, "country.name", "iso3c")) %>% 
  left_join(pop_iso) #join population data


```


```{r}
gap_22 <- happy_data %>% 
  left_join(pop_iso) %>% 
  filter(year == 2022)

happiest_country <- 
  gap_22 %>% 
  arrange(-happiness) %>% 
  head(1) %>% 
  pull(country)

happiest_country_value <- 
  gap_22 %>% 
  arrange(-happiness) %>% 
  head(1) %>% 
  pull(happiness)

lowest_happiness_country <- 
  gap_22 %>% 
  arrange(happiness) %>% 
  head(1) %>% 
  pull(country)

lowest_happiness_value <- 
  gap_22 %>% 
  arrange(happiness) %>% 
  head(1) %>% 
  pull(happiness)

average_happiness_value <-
  gap_22 %>%
  summarise(
    weighted_mean = sum(happiness * pop/ sum(pop, na.rm = TRUE), na.rm = TRUE)
  ) %>%
  round(1)

#correlate population size and child deaths!
```


```{r warning = FALSE}
country_shapes <- rnaturalearth::ne_countries()


# gap_map <-
#   left_join(country_shapes, gap_22,
#           by = c("adm0_a3" = "country_iso")) %>%
#   mutate(tooltip_label = paste(country,
#                                year,
#                                happiness,
#                                sep = ": ")) %>%
#   ggplot(aes(fill = happiness, text = tooltip_label, frame = year)) +
#   geom_sf()+
#   theme_void()+
#   theme(legend.position = "none")+
#   scale_fill_viridis_c()


#gap_map_ggplotly <- ggplotly(gap_map, tooltip = "text")

#using plot_ly here since geom_sf in combination with the "frame" option for the slider produced buggy output
gap_map_plot_ly <-
  left_join(country_shapes, happy_data,
            by = c("adm0_a3" = "country_iso")) %>%
  mutate(tooltip_label = paste0(country, ", ",
                                year, ": ",
                                happiness)) %>% 
  plot_ly(stroke = I("black"), split=~adm0_a3, color=~happiness, text=~tooltip_label, showlegend = FALSE, hoverinfo = "text", hoveron = "fills", frame=~year, alpha = 1) %>%
  style(hoverlabel = list(bgcolor = "white")) %>% 
  animation_slider(currentvalue = list(prefix = "Year ", font = list(color="#214773")))

```

```{r}
top_20_countries_2022 <- gap_22 %>%
  dplyr::arrange(desc(happiness)) %>%
  head(20) %>%
  mutate(tooltip_label = paste0(country, ", ",
                                year, ": ",
                                happiness)) %>%
  ggplot(aes(y = reorder(country, happiness),
             x = happiness,
             text = tooltip_label))+
  geom_col(fill = "#214773") +
  geom_text(aes(label = round(happiness, 1)),
            nudge_x = -10,
            color = "white"
            ) +
  labs(y = "Country",
       x = "Happiness [%]") +
  theme(legend.position = "none")+
  theme_minimal()

top_20_countries_ggplotly <-ggplotly(top_20_countries_2022, tooltip = "text")

```



# HOME

## Row 1 {height=25%}

```{r}
value_box(
  title = "Highest Happiness Score",
  value = happiest_country_value,
  showcase = bsicons::bs_icon("emoji-laughing"),
  theme = value_box_theme(bg = "#518fd6"),
  p(paste0("(", happiest_country, ")"))
)

# value_box(
#   title = div(class = "h2", "Highest Happiness Score"),  # Bootstrap h2 class for responsive text
#   value = div(class = "display-4", happiest_country_value),  # Bootstrap display-4 for larger text
#   showcase = bsicons::bs_icon("emoji-laughing"),
#   theme = value_box_theme(bg = "#518fd6"),
#   p(div(class = "h5", paste0("(", happiest_country, ")")))  # Bootstrap h5 class for smaller text
# )
```

```{r}
value_box(
  title = "Lowest Happiness score",
  value = lowest_happiness_value,
  showcase = bsicons::bs_icon("emoji-frown"),
  theme = value_box_theme(bg = "#214773"),
  p(paste0("(", lowest_happiness_country, ")"))
)
```

```{r}
value_box(
  title = "Average Happiness",
  value = average_happiness_value,
  showcase = bsicons::bs_icon("graph-up"),
  theme = value_box_theme(bg = "#3f71ab")
)
```


## Row 2 {height=75%}

### {width=70%}

```{r title = "Map of Happiness Score by Country over the Years"}
gap_map_plot_ly
```


### {width=30%}

```{r title = "Top 20 Countries by Happiness Score 2022"}
top_20_countries_ggplotly
```

# DOWNLOAD DATA

The data used in this dashboard is shown below and can be downloaded as a CSV. 

```{r}
library(htmltools)

htmltools::browsable(
  tagList(
    reactable(happy_data,
              elementId = "gapminder-table",
              searchable = T,
              filterable = T),

tags$button("Download as CSV",
            onclick = "Reactable.downloadDataCSV('gapminder-table')")
  )
)
```



# ABOUT

This data was downloaded from the website of the Gapminder Foundation.

Description: The happiness score is the national average response to the question of life evaluations asking the following "Please imagine a ladder, with steps numbered from O at the bottom to 10 at the top. The top of the ladder represents the best possible life for you and the bottom of the ladder represents the worst possible life
for you. On which step of the ladder would you say you personally feel you stand at this time?" This measure is also referred to as Cantril life ladder.

Gapminder has converted this indicator's scale from 0 to 100 to easly communicate it in terms of percentage.

Source_url: http://gapm.io/dhapiscore_whr

The Gapminder Foundation is a non-profit venture registered in Stockholm, Sweden, that promotes sustainable global development and achievement of the United Nations Millennium Development Goals by increased use and understanding of statistics and other information about social, economic, and environmental development at local, national, and global levels.

Gapminder was founded in 2005 by Ola Rosling, Anna Rosling Rönnlund, and Hans Rosling. The name Gapminder was derived from the "Mind the Gap" warning messages on the London Underground.

